CrypTool Script \\ Mathematics and Cryptography

### 2003-03-10 script/en/chapters/moderncryptography.tex
no message

\newpage
%**************************************************************************************************************************
\section{The Mathematical Ideas Behind Modern Cryptography}
 (Oyono R. / Esslinger B., September 2000, Update Nov. 2000)
\subsection{One way functions with trapdoor and complexity classes}
\index{Cryptography!modern} \index{One way function} \hypertarget{OneWayFunktion1}{}
A {\bf one way function} is a (one-to-one) function that can be calculated 
efficiently, but whose inverse is extremely complicated and practically 
impossible to calculate.\par

@@

\newpage
\subsection{Knapsack problem as a basis for public key procedures}
\index{Knapsack}
\subsubsection{Knapsack problem}

@@

\subsubsection{Merkle-Hellman knapsack encryption}
\index{Hellman Martin} \index{Merkle} In 1978, Merkle and Hellman 
\cite{Merkle1978} \index{Encryption!Merkle-Hellman} specified a public key 
encryption procedure that is based on ``defamiliarising'' the easy 0-1 
knapsack problem with a super-increasing sequence into a congruent one with a 
super-increasing sequence. It is a block ciphering that ciphers an $n$-bit 
plaintext each time it runs.
\index{Knapsack!Merkle-Hellman}
More precisely:

@@

%--------------------------------------------------------------------
\subsubsection[The RSA procedure]
{The RSA procedure\footnotemark}
\footnotetext{
Using CrypTool\index{CrypTool} v1.3 you gain practical experience
with the RSA procedure via the menu ``Indiv.Procedures \textbackslash 
RSA Demonstration \textbackslash RSA Cryptosystem''.
}
\index{RSA} \hypertarget{RSAVerfahren}{} 
As early as 1978, R. \index{Rivest Ronald} Rivest, \index{Shamir Adi} A. Shamir,  \index{Adleman Leonard} L. Adleman \cite{RSA1978} introduced 
the most important asymmetric cryptography procedure to date.  \par

@@

%--------------------------------------------------------------------
\subsubsection[Diffie-Hellman key agreement]
{Diffie-Hellman key agreement\footnotemark}
\footnotetext{%
With CrypTool\index{CrypTool} v1.3.04 you can execute the single steps of this 
exchange protocol with concrete numbers: \\
using menu {\bf Indiv. Procedures \textbackslash Diffie-Hellman Demonstration}.
}
\index{Key agreement (key exchange)!Diffie-Hellman} 
\index{Diffie Whitfield}

@@

    \subsubsection{ElGamal public key encryption procedure in $ \Z_p^\ast$}
\index{ElGamal!public key}
By varying the Diffie-Hellman key agreement protocol \index{Encryption!El Gamal public key} slightly, you can obtain an asymmetric encryption algorithm. This 
observation was made by Taher ElGamal.
\begin{center}
\fbox{\parbox{12cm}{        
Let $ p $ be a prime number such that the discrete logarithm in $ \Z_p $ is 
difficult to compute. Let $ \alpha \in \Z_p^\ast $ be a primitive element. Let $a \in \N$ and $ 
\beta = \alpha^{a}  \mod p. $\\
\underline{Public key:}
\[ p,\alpha,\beta. \]
\underline{Private key:}
\[a. \]
Let $ k \in \Z_{p-1} $ be a random number and $ x \in \Z_p^{\ast} $ the 
plaintext. \\
\underline{Encryption:}
\[ e_T(x,k)=(y_1,y_2), \]
where
\[ y_1=\alpha^k \mod p\]
and
\[ y_2 = x\beta^k \mod p.\]
\underline{Decryption:}
\[ d_T(y_1,y_2)= y_2 (y_1^{a})^{-1} \mod p \]
}}
\end{center}

@@

    \subsubsection{Generalised ElGamal public key encryption procedure}

@@

\newpage
\begin{thebibliography}{99}
\addcontentsline{toc}{subsection}{Bibliography}
           \bibitem[Adleman1982]{Adleman1982} Adleman L.: \index{Adleman Leonard}
           \emph{On breaking the iterated Merkle-Hellman public key 
Cryptosystem.} \\ 
           Advances in Cryptologie, Proceedings of Crypto 82, Plenum Press 1983, 
303-308.

 \bibitem[Balcazar1988]{Balcazar1988}\index{Balcazar 1988} Balcazar J.L., Daaz J., Gabarr\'{} J.: 
           \emph{Structural Complexity I.} \\ 
           Springer Verlag, pp 63.

       \bibitem[Brickell1985]{Brickell1985} Brickell E.F.:
           \emph{Breaking Iterated Knapsacks.}  \\
       Advances in Cryptology: Proc. CRYPTO\'84, Lecture Notes in Computer 
Science, vol. 196, 
           Springer-Verlag, New York, 1985, pp. 342-358.

           \bibitem[Lagarias1983]{Lagarias1983} Lagarias J.C.:
           \emph{Knapsack public key Cryptosystems and diophantine 
Approximation.} \\
           Advances in Cryptology, Proseedings of Crypto 83, Plenum Press.

           \bibitem[Merkle1978]{Merkle1978} Merkle R. and Hellman M.:
           \emph{Hiding information and signatures in trapdoor knapsacks.}  \\
           IEEE Trans. Information Theory, IT-24, 1978.

       \bibitem[RSA1978]{RSA1978} Rivest R.L., Shamir A. and Adleman L.:
           \emph{A Method for Obtaining Digital Signatures and Public Key 
Cryptosystems.} \\
           Commun. ACM, vol 21, April 1978, pp. 120-126.\index{RSA}

           \bibitem[Shamir1982]{Shamir1982} Shamir A.:
           \emph{A polynomial time algorithm for breaking the basic Merkle-
Hellman Cryptosystem.}\\
           Symposium on Foundations of Computer Science (1982), 145-152.
           \index{Shamir Adi}
%already defined in elementaryNumberTheory.inc
       \bibitem[Stinson1995]{4Stinson1995} Stinson D.R.:
           \emph{Cryptography.}  \index{Stinson 1995} \\
           CRC Press, Boca Raton, London, Tokyo, 1995.\index{Stinson 1995}
\end{thebibliography}
\section*{Web links} \addcontentsline{toc}{subsection}{Web links}
\begin{enumerate}
   \item 
\href{http://www.geocities.com/st_busygin/clipat.html}{http://www.geocities.com/st\_busygin/clipat.html }
\end{enumerate}

>>>>

\newpage
%**************************************************************************************************************************
\section{The Mathematical Ideas behind Modern Cryptography}
(Oyono R. / Esslinger B., Sep. 2000, Updates Nov. 2000, Feb. 2003)
\subsection{One way functions with trapdoor and complexity classes}
\index{Cryptography!modern} \index{One way function} \hypertarget{OneWayFunktion1}{}
A {\bf one way function} is a (one-to-one) function that can be calculated 
efficiently, but whose inverse is extremely complicated and practically 
impossible to calculate.\par

@@

\newpage
%--------------------------------------------------------------------
\subsection{Knapsack problem as a basis for public key procedures}%
\index{Cryptography!Public Key}
%--------------------------------------------------------------------
%--------------------------------------------------------------------
\subsubsection{Knapsack problem} \index{Knapsack}
%--------------------------------------------------------------------

@@

%--------------------------------------------------------------------
\subsubsection{Merkle-Hellman knapsack encryption}
%--------------------------------------------------------------------
\index{Hellman Martin} \index{Merkle Ralph}

In 1978, Merkle and Hellman 
\cite{Merkle1978} \index{Encryption!Merkle-Hellman} specified a public key 
encryption procedure that is based on ``defamiliarising'' the easy 0-1 
knapsack problem with a super-increasing sequence into a congruent one with a 
super-increasing sequence. It is a block ciphering that ciphers an $n$-bit 
plaintext each time it runs.
\index{Knapsack!Merkle-Hellman}
More precisely:

@@

%--------------------------------------------------------------------
\subsubsection[The RSA procedure]
{The RSA procedure\footnotemark}
\footnotetext{
Using CrypTool\index{CrypTool} v1.3 you gain practical experience
with the RSA procedure via the menu {\bf Indiv.Procedures \textbackslash{} 
RSA Demonstration \textbackslash{} RSA Cryptosystem}.
}
\index{RSA} \hypertarget{RSAVerfahren}{} 
As early as 1978, R. \index{Rivest Ronald} Rivest, \index{Shamir Adi} A. Shamir,  \index{Adleman Leonard} L. Adleman \cite{RSA1978} introduced 
the most important asymmetric cryptography procedure to date.  \par

@@

%--------------------------------------------------------------------
\subsubsection[Diffie-Hellman key agreement]
{Diffie-Hellman key agreement\footnotemark}
\footnotetext{%
With CrypTool\index{CrypTool} v1.3.04  this exchange protocol has been
visualized: you can execute the single steps with concrete numbers using 
menu {\bf Indiv. Procedures \textbackslash{} Diffie-Hellman Demonstration}.
}

\index{Key agreement (key exchange)!Diffie-Hellman} 
\index{Diffie Whitfield}
\index{Hellman Martin}
\index{Diffie-Hellman}

@@

%--------------------------------------------------------------------
\subsubsection{ElGamal public key encryption procedure in $ \Z_p^\ast$}
%--------------------------------------------------------------------
\index{ElGamal!public key}
By varying the Diffie-Hellman key agreement protocol \index{Diffie-Hellman} 
\index{Encryption!El Gamal public key} slightly, you can obtain an 
asymmetric encryption algorithm. This observation was made by Taher ElGamal.
\begin{center}
\fbox{\parbox{12cm}{        
Let $ p $ be a prime number such that the discrete logarithm in $ \Z_p $ is 
difficult to compute. Let $ \alpha \in \Z_p^\ast $ be a primitive element. Let $a \in \N$ and $ 
\beta = \alpha^{a}  \mod p. $\\
\underline{Public key:}
\[ p,\alpha,\beta. \]
\underline{Private key:}
\[a. \]
Let $ k \in \Z_{p-1} $ be a random number and $ x \in \Z_p^{\ast} $ the 
plaintext. \\
\underline{Encryption:}
\[ e_T(x,k)=(y_1,y_2), \]
where
\[ y_1=\alpha^k \mod p\]
and
\[ y_2 = x\beta^k \mod p.\]
\underline{Decryption:}
\[ d_T(y_1,y_2)= y_2 (y_1^{a})^{-1} \mod p \]
}}
\end{center}

@@

%--------------------------------------------------------------------
\subsubsection{Generalised ElGamal public key encryption procedure}
%--------------------------------------------------------------------

@@

%--------------------------------------------------------------------
\newpage
\begin{thebibliography}{99}
\addcontentsline{toc}{subsection}{Bibliography}

   \bibitem[Adleman1982]{Adleman1982} \index{Adleman 1982}
       Adleman L.: 
       \emph{On breaking the iterated Merkle-Hellman public key 
             Cryptosystem.} \\ 
       Advances in Cryptologie, Proceedings of Crypto 82, 
       Plenum Press 1983, 303-308.

   \bibitem[Balcazar1988]{Balcazar1988} \index{Balcazar 1988} 
       Balcazar J.L., Daaz J., Gabarr\'{} J.: 
       \emph{Structural Complexity I.} \\ 
       Springer Verlag, pp 63.

    \bibitem[Brickell1985]{Brickell1985} \index{Brickell 1985}
       Brickell E.F.:
       \emph{Breaking Iterated Knapsacks.}  \\
       Advances in Cryptology: Proc. CRYPTO\'84, 
       Lecture Notes in Computer Science, vol. 196, 
       Springer-Verlag, New York, 1985, pp. 342-358.

    \bibitem[Lagarias1983]{Lagarias1983} \index{Lagarias 1983} 
       Lagarias J.C.:
       \emph{Knapsack public key Cryptosystems and diophantine
             Approximation.} \\
       Advances in Cryptology, Proseedings of Crypto 83, Plenum Press.

    \bibitem[Merkle1978]{Merkle1978} \index{Merkle 1978} 
       Merkle R. and Hellman M.:
       \emph{Hiding information and signatures in trapdoor knapsacks.} \\
       IEEE Trans. Information Theory, IT-24, 1978.

    \bibitem[RSA1978]{RSA1978} \index{RSA 1978}
       Rivest R.L., Shamir A. and Adleman L.:
       \emph{A Method for Obtaining Digital Signatures and 
             Public Key Cryptosystems.} \\
       Commun. ACM, vol 21, April 1978, pp. 120-126.

    \bibitem[Shamir1982]{Shamir1982} \index{Shamir 1982}
       Shamir A.:
       \emph{A polynomial time algorithm for breaking the basic 
             Merkle-Hellman Cryptosystem.} \\
       Symposium on Foundations of Computer Science (1982), 145-152.

%already defined in elementaryNumberTheory.tex
    \bibitem[Stinson1995]{4Stinson1995} \index{Stinson 1995}
       Stinson D.R.:
       \emph{Cryptography.} \\
       CRC Press, Boca Raton, London, Tokyo, 1995.

\end{thebibliography}

%--------------------------------------------------------------------
\section*{Web links} \addcontentsline{toc}{subsection}{Web links}
\begin{enumerate}
   \item \href{http://www.geocities.com/st_busygin/clipat.html}
              {http://www.geocities.com/st\_busygin/clipat.html }
\end{enumerate}
